**Dimensionality reduction techniques**

```{r}
library(MASS)
human = read.table("data/human.txt", sep="\t")
dim(human)
```
Details can be seen on the Human Development Index site. The dataset has 155 observations of 8 variables.

```{r}
library(GGally)
ggpairs(human)
```
Plenty of correlations - for instance, the maternal mortality rate and life expectancy are highly negatively correlated (-0.857) while expected education in years is highly positively correlated to life expectancy (0.789). There not so much women in the parliament. There are more women with education rather with employment. 

```{r}
pca_human <- prcomp(human)
summary(pca_human)
pca_pr <- round(100*summary(pca_human)$importance[2, ], digits = 1)
pc_lab <- paste0(names(pca_pr), " (", pca_pr, "%)")
biplot(pca_human, choices = 1:2,cex = c(0.5, 0.6),xlab = pc_lab[1], ylab = pc_lab[2])
```
GNI highly influence the first component of PCA, and so it captures almost all the variability, perhaps due to the unstandardized data.

```{r}
human_std <- scale(human)
pca_humanstd <- prcomp(human_std)
pca_pr <- round(100*summary(pca_humanstd)$importance[2, ], digits = 1)
pc_lab <- paste0(names(pca_pr), " (", pca_pr, "%)")
biplot(pca_humanstd, choices = 1:2,cex = c(0.5, 0.6),xlab = pc_lab[1], ylab = pc_lab[2])
```

Health and education variables contribute more to the first component, while labour and parliament variables - to the second. Standardization helped with the disproportions of component's contribution - now the first component explains around half of variance, and together with the second they explain around 70 percentage of it.

```{r}
library(FactoMineR)
library(dplyr)
library(tidyr)
library(ggplot2)
data("tea")
str(tea)
summary(tea)
keep_columns <- c("Tea", "How", "how", "sugar", "where", "lunch")
tea_time <- dplyr::select(tea, one_of(keep_columns))
tidyr::gather(tea_time) %>% ggplot(aes(value)) + facet_wrap("key", scales = "free")+geom_bar()+theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))
summary(tea_time)
str(tea_time)
```

The tea dataset contains 300 observations and 36 variables.
Tea is drinked separately most usually separatedly from any additions, not in the lunch time, and simply using tea bag from chain store. The distribution of sugar/no sugar is almost unimodal.

```{r}

mca <- MCA(tea_time, graph = FALSE)
summary(mca)
plot(mca, invisible=c("ind"))
```

It make sense that unpackaged tea is near the tea shop while tea bag is near chain store. Also it seems that there is a distinction between "black tea with limon, no sugar" and "Earl Grey with milk and sugar" - fast googling shows that these combinations are typical.